---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# safar6

<!-- badges: start -->
[![Project Status: Concept – Minimal or no implementation has been done
yet, or the repository is only intended to be a limited example, demo,
or
proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
[![R-CMD-check](https://github.com/matt-dray/safar6/workflows/R-CMD-check/badge.svg)](https://github.com/matt-dray/safar6/actions)
[![CodeFactor](https://www.codefactor.io/repository/github/matt-dray/safar6/badge)](https://www.codefactor.io/repository/github/matt-dray/safar6)
<!-- badges: end -->

## Summary 

Work in progress: an R package containing an [R6](https://r6.r-lib.org/) class to simulate simplified gameplay elements of [the Safari Zone](https://bulbapedia.bulbagarden.net/wiki/Kanto_Safari_Zone) sub-area from [_Pokémon Blue_](https://bulbapedia.bulbagarden.net/wiki/Generation_I) (1998) with OOP.

The first generation of _Pokémon_ games were developed for [the Nintendo Game Boy](https://en.wikipedia.org/wiki/Game_Boy) by [Game Freak](https://www.gamefreak.co.jp/) and published by [Nintendo](https://www.nintendo.com). Pokémon as a property is owned by [The Pokémon Company](https://www.pokemon.co.jp/).

This project is a demonstration of [the {R6} package](https://r6.r-lib.org/) and the R6-class for object-oriented programming in R. It's purely for learning purposes.

## Basic gameplay

Professor Oak:

> `Your very own POKéMON legend is about to unfold! A world of dreams and adventures with POKéMON awaits! Let's go!`

Install from GitHub with `remotes::install_github("matt-dray/safar6")`.

The game is text-based and takes place entirely in the console. Start a game like `sz <- safar6::SafariZone$new()`. Be warned: it's tricky. But then so was the original.

Use `sz$step()` to take a step through the tall grass of the Safari Zone. There's a chance you'll encounter a wild Pokémon; perhaps it will be a rare one at a high level. You'll be prompted to throw a Safari Ball to try and catch it, with the option to throw bait or a rock to influence its catchability. Try to catch as many Pokémon as you can before you run out of steps or balls.

Use `sz$pause()` to 'pause' the game and see your current stats. You can boot up `$bills_pc` to see which Pokémon you've caught so far. 

For help with fields and methods see `?safar6::SafariZone`.

<details><summary>Click for (contrived!) example gameplay `r emo::ji("space_invader")` </summary>
```{r, eval=FALSE}
> sz <- safar6::SafariZone$new()
# Welcome to the SAFARI ZONE!
#  For just $500, you can catch all the Pokemon you want in the park!
#  Would you like to join the hunt?
#  > YES NO
#  That'll be $500 please!
#  We only use a special POKe BALL here.
#  BLUE received 30 SAFARI BALLs!
#  We'll call you on the PA when you run out of time or SAFARI BALLs!
> sz$step()
# 499/500
> sz$step()
# 498/500
> sz$step()
# 497/500
# Wild VENONAT L22 appeared!
# ------------------------
# BALLx30 (1)     BAIT (2)
# THROW ROCK (3)  RUN (4)
# Selection: 
> 3
# BLUE threw a ROCK.
# Wild VENONAT is angry!
# ------------------------
# BALLx30 (1)     BAIT (2)
# THROW ROCK (3)  RUN (4)
# Selection:
> 1
# BLUE used SAFARI BALL!
# Wobble...
# Darn! The POKeMON broke free!
# Wild VENONAT is angry!
# ------------------------
# BALLx29 (1)     BAIT (2)
# THROW ROCK (3)  RUN (4)
# Selection: 
> 1
# BLUE used SAFARI BALL!
# Wobble... Wobble... Wobble...
# All right!
# VENONAT was caught!
# Do you want to give a nickname to VENONAT?
# Selection (YES/NO):
> YES
# Nickname: 
> TAJIRI
# TAJIRI was transferred to BILL's PC!
> sz$pause()
# 497/500
# BALLx28
# Transferred to BILL's PC: 1
> sz$bills_pc
#   nickname species level
# 1   TAJIRI RHYHORN    25
```
</details>

## Details

### The Safari Zone

The [Safari Zone](https://bulbapedia.bulbagarden.net/wiki/Kanto_Safari_Zone) is an enclosed area separate from the rest of the game. Game mechanics are different here: you're allowed 500 steps and 30 balls to capture Pokémon. Also, wild encounters don't result in a 'battle'; instead you can 'bait' or 'throw a rock' to influence the catch rate of the target. I chose to simulate this area due to its relative simplicity compared to other areas of the game, which use a more complex battle mechanic.

### Technicals

The fields in the `SafariZone` class are 'overworld' counters, basically: the number of steps and balls remaining, and the number and identity of captures you've made.

Methods are:

* `$new()` to initialise the class and start the game
* `$step()`, which moves your 'player' and may result in an encounter
* `$pause()` (a copy of the `$print()` method) to simulate the pause screen from the game, showing the current number of steps and balls remaining

The `$step()` method does the hard work. It contains code that:

* decides whether you encountered a target, based on the encounter rate of the Safari Zone (~12%)
* selects a wild Pokémon based on the relative encounter-rate slots for the Safari Zone Center Area particular to _Pokémon Blue_ (e.g. 20% for a level 22 Nidoran Female through to 1% for a level 23 Chansey; see `safar6::pokemon` for details)
* generates individual variation in the selected Pokémon's speed and HP (IV values), which influence catchability
* prompts you interactively for an action: throw ball, bait, throw rock, run
* adjusts the 'angry' and 'eating' values of the target and modifies catch rate and run probability as a result (a rock doubles the catch rate and the flee chance; bait halves the catch rate, but makes it four times less likely to run)
* calculates whether the Pokémon will be caught or will run away
* terminates the game if steps or balls are depleted

## Development

### TODO

* "First, what is your name?"
* "New POKeDEX data will be added for [x]."
* "Do you want to give a nickname to [x]?"
* an Easter egg...

Feel free to add ideas as [issues](https://www.github.com/matt-dray/safar6/issues/).

### Not included

To keep things (much!) simpler than the original, this implementation doesn't include a number of features from the original game. Some missing elements are:

* visuals (everything is printed in the console)
* movement around a map (all steps are treated as though you're moving through tall grass)
* all areas of the Safari Zone (only the Center area is used here)
* finding items (they have no use in this context)
* fishing and use of different rod types
* species catch and encounter rates from the other Gen 1 versions (e.g. Scyther is in _Red_ rather than Pinsir; and _Yellow_ has some catch-rate changes)

## Sources

Information about game mechanics and values are relatively tricky to come by. I used the following resources:

* [Bulbapedia](https://bulbapedia.bulbagarden.net/) for Pokémon stats and game mechanics
* [The Cave of Dragonflies](https://www.dragonflycave.com/) for game mechanics
* the [Pokémon Slots](https://sites.google.com/site/pokemonslots) website for encounter rates
* the [pret/pokered GitHub repo](https://github.com/pret/pokered) for a disassembly of the games

I also referred to a post I wrote about [using {R6} to simulate an Automatic Bell Dispenser](https://www.rostrum.blog/2020/04/04/repaying-tom-nook-with-r6/) from Nintendo's _Animal Crossing: New Horizons_ (2020).

## Code of Conduct

Please note that the {safar6} project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

## Disclaimer

I do not condone throwing rocks at any creature, let alone catching and imprisoning them.
